---
title: "Data Source"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set( root.dir = "~/Desktop/data_souce/Nov2017_hana/")
source("~/Packaging_Type/_R/Function.R")
```

## Source

### [GISAID](https://www.gisaid.org)


#### Searching criteria

<font size="2">

Type | H | Host | Location | Required Segments | only com. Min Len. | only GISAID 
---- | - | ---- | -------- | ----------------- | -------------------| ----------- 
A | 5 | __w/o__( Lab derived, Unknown, NA ) | all | HA + NA | 1000 ( __w/o__ check ) | check 

</font>
<br>

#### FASTA header

```{r, eval = FALSE}
Isolate name Type Collection date Isolate ID
```

<br>

#### Raw file

1. ##### `pH5_G_2136_20171124.fasta`
2. ##### `pNA_G_2135_20171124.fasta`
3. ##### `G_2135_20171124.xls`
4. ##### `ac_20171124.xls`

<br>

### NCBI

#### Searching criteria

<font size="2">

Type | Host | Country | Segment | Subtype | Req. segments | pH1N1 | Lab | lineage def. | FLU | Vx | Mix 
---- | ---- | ------- | ------- | ------- | ------------- | ----- | --- | ------------ | --- | -- | --- 
A | __w/o__( Unknown ) | any | HA or NA | 5 | 4 + 6 | E | E | I | I | I | E 

</font>

linkage: 
[HA](https://www.ncbi.nlm.nih.gov/genomes/FLU/Database/nph-select.cgi?aln_subtype=5&cmd=show_query&country=any&defline_saved=%3E%7Baccession%7D%20%7Bstrain%7D%20%7Byear%7D/%7Bmonth%7D/%7Bday%7D%20%7Bsegname%7D&go=database&host=Avian&host=Bat&host=Blow%20fly&host=Bovine&host=Camel&host=Canine&host=Cat&host=Cheetah&host=Civet&host=Environment&host=Equine&host=Feline&host=Ferret&host=Giant%20anteater&host=Human&host=Leopard&host=Lion&host=Mink&host=Mouse&host=Muskrat&host=Panda&host=Penguin&host=Pika&host=Raccoon%20dog&host=Reassortant&host=Seal&host=Sloth%20bear&host=Stone%20marten&host=Swine&host=Tiger&host=Whale&host=Wild%20boar&lab=exclude&lineage=include&niaid=include&orderby=date%20asc&prev_cmd=show_query&reqseg=4&reqseg=6&searchin=strain&segment=4&sequence=N&showfilters=true&subtype_h=5&subtype_mix=exclude&subtype_n=any&swine=exclude&taxid=197911&type=a&vac_strain=include)
[NA](https://www.ncbi.nlm.nih.gov/genomes/FLU/Database/nph-select.cgi?aln_subtype=5&cmd=show_query&country=any&defline_saved=%3E%7Baccession%7D%20%7Bstrain%7D%20%7Byear%7D/%7Bmonth%7D/%7Bday%7D%20%7Bsegname%7D&go=database&host=Avian&host=Bat&host=Blow%20fly&host=Bovine&host=Camel&host=Canine&host=Cat&host=Cheetah&host=Civet&host=Environment&host=Equine&host=Feline&host=Ferret&host=Giant%20anteater&host=Human&host=Leopard&host=Lion&host=Mink&host=Mouse&host=Muskrat&host=Panda&host=Penguin&host=Pika&host=Raccoon%20dog&host=Reassortant&host=Seal&host=Sloth%20bear&host=Stone%20marten&host=Swine&host=Tiger&host=Whale&host=Wild%20boar&lab=exclude&lineage=include&niaid=include&orderby=date%20asc&prev_cmd=show_query&reqseg=4&reqseg=6&searchin=strain&segment=6&sequence=N&showfilters=true&subtype_h=5&subtype_mix=exclude&subtype_n=any&swine=exclude&taxid=197911&type=a&vac_strain=include)

<br>

#### FASTA header

```{r, eval = FALSE}
>{accession}_{strain}_{serotype}_|{country}|_{year}-{month}-{day}
```

<br>

#### Raw file

1. ##### `pH5_N_5753_20171124.fasta`
2. ##### `pNA_G_2135_20171124.fasta`

***

## Pre-process

1. Save `G_2135_20171124.xls` as `G_2135_20171124.csv ` . 
2. Replace _blank_ with _undersocre_ in NCBI filess. 

```{r}
system( 'cd ./Pre_process; for f in *.fasta; do sed -i "" "s/ /_/g" $f; done' )
```

***
## Sequence preparation and curation 

The process includes eliminating duplicated sequences, extracting isolatation info. from names, matching HA-NA data and sequence curation. More detail in `pH5_seqPrep.R`.



***

## Selection 

### remove re-introduction transmission

1. Manually map trees with __China__ and __Hong_Kong__ sequence on reference (big) trees.
2. Record strains emerged in other countries.
3. Label them on the samll tree. 

* #### clade 2.3.4 

```{r, eval = FALSE}
CY098695_Guangxi_1_2008_|China|_H5N1_2008.115
CY098730_Xinjiang_1_2009_|China|_H5N1_2009.025
CY098830_water_Xinjiang_3_2009_|China|_H5N1_2009.041
CY098838_environment_Xinjiang_6_2009_|China|_H5N1_2009.041
EPI135216_Guizhou_1_2013_|China|_H5N1_2013.104
EPI135217_Guizhou_2_2013_|China|_H5N1_2013.107
EPI102993_Guizhou_1_2012_|China|_H5N1_2012.058
CY098798_environment_Guizhou_2_2009_|China|_H5N1_2009.041
CY098808_environment_Guizhou_4_2009_|China|_H5N1_2009.041
```

* #### clade 2.3.2

```{r, eval = FALSE}
KX247924_environment_Changsha_115_2014_|China|_H5N1_2014.189
KX247927_environment_Changsha_116_2014_|China|_H5N1_2014.189
EPI219772_Environment_Hunan_15252_2014_|China|_H5N1_2014.112
KX247933_environment_Changsha_213_2014_|China|_H5N1_2014.285
KX247918_environment_Changsha_15_2014_|China|_H5N1_2014.104
KX247915_environment_Changsha_14_2014_|China|_H5N1_2014.104
KX247921_environment_Changsha_42_2014_|China|_H5N1_2014.110
EPI219757_Environment_Guangxi_14313_2013_|China|_H5N1_2013.923
KX247936_environment_Changsha_155_2014_|China|_H5N1_2014.252
CY146644_duck_Hunan_S4030_2011_|China|_H5N1_2011.855
JX576788_environment_Hunan_3_2011_|China|_H5N1_2011.178
EPI76081_Hubei_1_2010_|China|_H5N1_2010.414
MF116311_duck_China_E53_2012_|China|_H5N1_2012.381
MF116310_goose_China_E43_2012_|China|_H5N1_2012.381
EPI201002_Black_necked_Grebe_Inner_Mongolia_Brain2_2015_H5N1_|China|_H5N1_2015.359
EPI201003_Black_necked_Grebe_Inner_Mongolia_Lung2_2015_H5N1_|China|_H5N1_2015.359
EPI201004_Black_necked_Grebe_Inner_Mongolia_Lung3_2015_H5N1_|China|_H5N1_2015.359
EPI201005_Black_necked_Grebe_Inner_Mongolia_Liver3_2015_H5N1_|China|_H5N1_2015.359
EPI201006_Black_necked_Grebe_Inner_Mongolia_Brain3_2015_H5N1_|China|_H5N1_2015.359
EPI201007_Black_necked_Grebe_Inner_Mongolia_Lung3_2_2015_H5N1_|China|_H5N1_2015.359
EPI201008_Black_necked_Grebe_Inner_Mongolia_Oropharynx_swab3_2015_H5N1_|China|_H5N1_2015.359
EPI201009_Black_necked_Grebe_Inner_Mongolia_Trachea3_2015_H5N1_|China|_H5N1_2015.359
EPI201001_Black_necked_Grebe_Inner_Mongolia_Brain1_2015_H5N1_|China|_H5N1_2015.359
```

### remove reassortants

1. Create __cleantree__ by removing the accession number
2. Examine the reassortment with tangleplot 
3. Select the region of nonreassorted lineage by defining the __y-axis__ of HA & NA tree
4. Extract the sequence based on the eligibel HA-NA pairs

Details: `pH5_seqPrep.R`

















